The finite representation property for some reducts of relation algebras

### 2021-03-13 frp-paper/Paper.tex
[Problem-5.18] the definition of a network and a game
important fixes in frp-paper/Paper.tex

We also show that the class of representable join semilattice-ordered semigroups has a recursively
enumerable recursive axiomatisation and the finite representation property.
\end{abstract}

@@

We will use the following notation due to, for example, \cite{hirsch2011positive}. Let $\tau$ be a subset of operations and predicates definable in ${\bf RA}$. ${\bf R}(\tau)$ is the class of subalgebras of $\tau$-subreducts of algebras belonging to ${\bf RRA}$. We assume that ${\bf R}(\tau)$ is closed under isomorphic copies.

@@

A quantale is a complete lattice-ordered semigroup. Quantales has been initially introduced by Mulvey to provide a noncommutative generalisation of locales, see \cite{mulvey1986suppl}. The rigorous definition is the following:

@@

As a result, we have the following representation theorem.

@@

  A prenetwork over $\mathcal{A} = (V, E, l)$ is a \emph{network} if the following hold:
  \begin{enumerate}
    \item For every $u, v \in V$ and for every $x,y,z \in \mathcal{A}$ $z \in l(u, v)$ and $z \leq x \: ; \: y$ implies $x \in N(u, w)$ and $y \in l(w, v)$ for some $w \in V$.
    \item For every $u, v, w \in V$ one has $l(u, v) ; l(v, w) \subseteq l(u, w)$.
    \item For every $u, v \in V$ $l(u,v)$ is join-prime. That is, for each $a, b \in \mathcal{A}$ if $a + b \in l(u,v)$, then either $a \in l(u,v)$ or $b \in l(u,v)$.
  \end{enumerate}
\end{definition}

@@

    Let us define the $rep : \mathcal{A} \to \operatorname{Nodes}(\mathcal{N})$:
    \begin{center}
      $rep(a) = \{ (x, y) \: | \: \exists b \in \down{a} \:\: b \in l_{\mathcal{N}}(x, y)\}$
    \end{center}

    Let us ensure that $rep$ is a representation.

    \begin{itemize}
  \item Let us show that $rep(a + b) = rep(a) \cup rep(b)$

  Suppose $(x, y) \in rep(a + b)$. That is, there exists $c \leq a + b$ with $c \in l_{\mathcal{N}}(x, y)$, so does $a + b$ since $l_{\mathcal{N}}$ is an upper cone.

@@

That is, there are $a_1, a_0 \in \mathcal{A}$ with $a_1 \not\leq a_0$ such that $a + b \in l_{\mathcal{N}^{*}(a_1, a_0)}(x,y)$.
This means $a + b \in \bigcup \limits_{i < \omega} \mathcal{N}_i$.
$\forall$ makes all possible join moves, so either $a \in l_{\mathcal{N}_{i + 1}}(x, y)$ or $b \in l_{\mathcal{N}_{i + 1}}(x, y)$ depending on what $\exists$ chooses.
Thus, $(x, y) \in rep(a) \cup rep(b)$.

Suppose for the converse, $(x, y) \in rep(a)$. We need $(x, y) \in rep(a + b)$.
In other words, we have some $c \in \mathcal{A}$ with $c \leq a$ and $c \in l_{\mathcal{N}}(x, y)$.
We have $c \leq a \leq a + b$, so $(x, y) \in rep(a + b)$.

\item Let us show that $rep(a ; b) = rep(a) ; rep(b)$.

Suppose $(x, y) \in rep(a ; b)$. We need some $z$ with $(x, z) \in rep(a)$ and $(z, y) \in rep(b)$.

@@

By the condition, $\forall$ makes the witness moves and $\exists$ responds with a witness. Her response is a fresh node $z$ such that $l_{\mathcal{N}_{i + 1}}(x, z) =  \up{(l_{\mathcal{N}_i}(x, z) \cup \{ a \})}$ and $l_{\mathcal{N}_{i + 1}}(z, y) = \up{(l_{\mathcal{N}_i}(z, y) \cup \{ b\})}$. The inclusion $rep(a ; b) \subseteq rep(a) ; rep(b)$ holds since all witness moves have been played and for every such a move there exists a witness.

\item Suppose $(x, y) \in rep(a) ; rep(b)$. We need $(x, y) \in rep(a;b)$. That is, there exists $z \in \operatorname{Nodes}(\mathcal{N})$ with $(x, z) \in rep(a)$ and $(z, y) \in rep(b)$. So, there are $c, d$ such that $c \leq a$ with $c \in l_{\mathcal{N}}(x, z)$ and $d \leq b$ with $d \in l_{\mathcal{N}}(z, y)$.
We also know that $l_{\mathcal{N}}(x, z) ; l_{\mathcal{N}}(z, y) \subseteq l_{\mathcal{N}}(x, y)$ because all composition moves have been played.
So $c ; d \in l_{\mathcal{N}}(x, y)$. That makes $(x, y) \in rep(a;b)$ since $c ; d \leq a ; b$.

@@

Let us define the equivalence relation $\approx$ on $T$ as follows for $t_1$ and $t_2$ are terms:
\begin{center}
$t_1 \approx t_2$ iff $F(\omega) \models t_1 = t_2$.
\end{center}
where $F(\omega)$ is the free countably generated algebra join semilattice-ordered semigroup.
We have the following partial order on $T$ defined as
\begin{center}
  $t_1 \preceq t_2$ iff $t_1 + t_2 \approx t_2$, that is, $F(\omega) \models t_1 + t_2 = t_2$
\end{center}

@@

Let $\mathcal{N}$ be a term network. We define the following three extensions of $\mathcal{N}$ reflecting the composition, witness, and join moves respectively:
\begin{enumerate}
  \item Let $x, y \in \operatorname{Nodes}(\mathcal{N})$ and $t$ a term. $\mathcal{N}_c$ is the extension of $\mathcal{N}$,
  where $\operatorname{Nodes}(\mathcal{N}_c) = \operatorname{Nodes}(\mathcal{N})$ and $\up{l_{\mathcal{N}}(x, y) \cup t}$ and $l_{\mathcal{N}_c}(u, v) = l_{\mathcal{N}}(u, v)$ for each $u \neq x$ and $v \neq y$. We denote this network as $\mathcal{N}_c(\mathcal{N}, x, y, t)$.
  \item
  Let $x, y \in \operatorname{Nodes}(\mathcal{N})$, $z$ a fresh node, and $t_1$, $t_2$ any terms. Let us define a network $T$ such that $\operatorname{Nodes}(T) = \{ x, y, z\}$. We define labelling as $l_{T}(x, y) = \up{t_1}$ and $l_{T}(y, z) = \up{t_2}$. So we put $\mathcal{N}_w = \mathcal{N} \cup T$.
  We denote this network as $\mathcal{N}_w(\mathcal{N}, x,y,z, t_1, t_2)$.
  \item Let $x, y \in \operatorname{Nodes}(\mathcal{N})$ and let $t_1, t_2$ be terms
  Let us define $T_i = \langle \{ x, y\}, \{ (x, y) \}, l_{T_i} \rangle$, where $l_{T_i}(x, y) = \up{(l_{\mathcal{N}}(x, y) \cup t_i)}$ for $i = 1,2$. So $\mathcal{N}_{j_1} = \mathcal{N} \cup T_1$ and $\mathcal{N}_{j_2} = \mathcal{N} \cup T_2$.
  We denote this as $\mathcal{N}_j(\mathcal{N}, x, y, t_1, t_2)$.
\end{enumerate}

@@

One may consider $\mathcal{A}$, a finite join semilattice-ordered semigroup, as a quantale since the underlying semilattice is complete being finite. By Theorem~\label{quantaleRep}, $\mathcal{A}$ is isomorphic to the relational quantale $\widehat{A}$. From the other hand, if $\mathcal{Q}$ is a relational quantale, then $\Sigma A \subseteq \bigcup A$ for $A \subseteq \mathcal{Q}$. The converse inclusion does not hold in general, see \cite[Remark 3.2 and Example 3.6]{brown1993representation}.

@@

A quantale $\mathcal{Q}$ is called \emph{simple}, if
$\sum \limits_{i \in I} \widehat{A} = \bigcup \limits_{i \in A} \widehat{A}$ for $A \subseteq \mathcal{Q}$.

@@

\begin{proposition} \label{distr}
  Let $\mathcal{A} \in {\bf R}(;,+)$ and $\mathcal{A}$ is finite, then
  \begin{enumerate}
    \item $\mathcal{A}$ is distributive,
    \item Every generator of $\mathcal{A}$ as a quantale is join-prime.
  \end{enumerate}
\end{proposition}

@@

  \begin{enumerate}
    \item
    $\mathcal{A}$ is finite, so the $\operatorname{Id}(\mathcal{A})$ consists of principal ideals, that is, lower cones generated by elements of $\mathcal{A}$.

    \begin{center}
    $\operatorname{Id}(\mathcal{A}) = \{ \down{a} \: | \: a \in \mathcal{A} \}$
    \end{center}

    Let $h$ be a representation of $\mathcal{A}$, consider $h(\mathcal{A})$ and its set of ideals $\operatorname{Id}(h(\mathcal{A}))$. Intersection of ideals is defined is usual since intersection of ideals is an ideal, join is the defined as follows:

    \begin{center}
    $\down{h(a)} \vee \down{h(b)} = \down{(h(a) \cup h(b))}$
    \end{center}
    Note that $\down{(h(a) \cup h(b))} = \down{h(a + b)}$ since $h$ is a representation of $\mathcal{A}$.

@@

    Suppose $h(d) \in \down{h(c)} \cap \down{h(a + b)}$.
    $h(d) \in \down{h(a + b}$ implies $h(d) \subseteq h(a + b)$.

    Then if $(x, y) \in h(d)$, then either $(x, y) \in h(a)$ or $(x, y) \in h(b)$.
    If $(x, y) \in h(a)$ (or $h(b)$), then we have $h(d) \in \down{(h(c) \cap h(a))}$ (or $\down{(h(c) \cap h(b))}$).

    Suppose $h(d) \in (\down{h(c)} \cap \down{h(a)}) \vee (\down{h(c)} \cap \down{h(b)})$. So, we have either $h(d) \in \down{h(c)} \cap \down{h(a)}$ or $h(d) \in \down{h(c)} \cap \down{h(b)}$. So $h(d) \in \down{h(c)}$ anyway. If $h(d) \in \down{h(a)}$, then $d \leq a \leq a + b$, so $h(d) \in \down{h(a + b)}$.
    The second alternative is the same as the first one.

    So, $\operatorname{Id}(h(\mathcal{A}))$ is distributive, and, therefore, $\mathcal{A}$ is distibutive as a join-semilattice by \cite[Lemma 184]{gratzer2011lattice}.

    \item Let $g \in \mathcal{A}$ be a generator, suppose $g \leq a + b$, so there are $a' \leq a$ and $b' \leq b$ with $g = a' + b'$. $g$ is join-irreducible, thus, one has either $g = a'$ or $g = b'$. In fact, this statement generalises the fact that an element is join-prime iff it is join-irreducible in distributive lattices, see \cite[Lemma 5.11]{davey2002introduction}.
  \end{enumerate}
\end{proof}

The following lemma allows one to consider finite elements of ${\bf R}(;,+)$ as simple quantales. If $\mathcal{A}$ is finite and representable, then every $a$ has the form $\Sigma \{ g \in \mathcal{G}(\mathcal{A}) \: | \: g \leq a \}$, but every generator is join-prime by Proposition~\ref{distr}. So we have:

\begin{proposition}\label{quantalerepisok}
  Let $\mathcal{A}$ be a finite representable join-semilattice semigroup, $\mathcal{A}$ is completely coprime algebraic as a quantale and, therefore, is simple.
\end{proposition}

@@

\begin{proof}
  Let us show that if $\mathcal{A} \in {\bf R}(+, ;)$ such that $\mathcal{A}$ is finite, then
  $h(\mathcal{A})$ is isomorphic to some subalgebra of $\widehat{A}$. Consider the map
  $f : h(a) \mapsto \hat{a}$. $f$ preserves composition since $h$ and $\hat{.}$ are semigroup homomorphisms. The equality $f(h(a) \cup h(b)) = \hat{a} \cup \hat{b}$ holds since $\mathcal{A}$ is simple as a quantale by Proposition~\ref{quantalerepisok}.

  $\mathcal{A}$ is a relational quantale on the underlying of $\mathcal{A}$, but $\mathcal{A}$ is finite, so $\mathcal{A}$ has a representation over a finite base.
\end{proof}

@@

The author is sincerely grateful to Robin Hirsch, Ian Hodkinson, and his supervisor Ilya Shapirovsky for valuable comments. The previous version of this paper contained an error. Many thanks to Jas Semrl for noticing that.

>>>>

We also show that the class of representable join semilattice-ordered semigroups has a recursively
enumerable axiomatisation and the finite representation property.
\end{abstract}

@@

Let $\tau$ be a subset of operations and predicates definable in ${\bf RA}$. ${\bf R}(\tau)$ is the class of subalgebras of $\tau$-subreducts of algebras belonging to ${\bf RRA}$. We assume that ${\bf R}(\tau)$ is closed under isomorphic copies.

@@

A quantale is a complete lattice-ordered semigroup. Quantales has been initially introduced by Mulvey to provide a noncommutative generalisation of locales, see \cite{mulvey1986suppl}.

@@

As a result, we have the following representation theorem \cite[Corollary 2]{goldblatt2006kripke}.

@@

  A prenetwork over $\mathcal{A} = (V, E, l)$ is a \emph{network} if the following hold:
  \begin{enumerate}
    \item ({\bf Saturation condition}) For every $u, v \in V$ and for every $x,y,z \in \mathcal{A}$ $z \in l(u, v)$ and $z \leq x \: ; \: y$ implies $x \in N(u, w)$ and $y \in l(w, v)$ for some $w \in V$.
    \item ({\bf Coherence conidition}) For every $u, v, w \in V$ one has $l(u, v) ; l(v, w) \subseteq l(u, w)$.
    \item ({\bf Join-primeness}) For every $u, v \in V$ $l(u,v)$ is join-prime. That is, for each $a, b \in \mathcal{A}$ if $a + b \in l(u,v)$, then either $a \in l(u,v)$ or $b \in l(u,v)$.
  \end{enumerate}
\end{definition}

@@

    Let us define the $\operatorname{rep} : \mathcal{A} \to \operatorname{Nodes}(\mathcal{N})$:
    \begin{center}
      $\operatorname{rep}(a) = \{ (x, y) \: | \: \exists b \in \down{a} \:\: b \in l_{\mathcal{N}}(x, y)\}$
    \end{center}

    Let us ensure that $\operatorname{rep}$ is a representation.

    \begin{itemize}
  \item Let us show that $\operatorname{rep}(a + b) = \operatorname{rep}(a) \cup \operatorname{rep}(b)$

  Suppose $(x, y) \in \operatorname{rep}(a + b)$. That is, there exists $c \leq a + b$ with $c \in l_{\mathcal{N}}(x, y)$, so does $a + b$ since $l_{\mathcal{N}}$ is an upper cone.

@@

That is, there are $a_1, a_0 \in \mathcal{A}$ with $a_1 \not\leq a_0$ such that $a + b \in l_{\mathcal{N}^{*}(a_1, a_0)}(x,y)$.
This means $a + b \in \bigcup \limits_{i < \omega} \mathcal{N}_i$.
$\forall$ makes all possible join moves, so either $a \in l_{\mathcal{N}_{i + 1}}(x, y)$ or $b \in l_{\mathcal{N}_{i + 1}}(x, y)$ depending on what $\exists$ chooses.
Thus, $(x, y) \in \operatorname{rep}(a) \cup \operatorname{rep}(b)$.

Suppose for the converse, $(x, y) \in \operatorname{rep}(a)$. We need $(x, y) \in \operatorname{rep}(a + b)$.
In other words, we have some $c \in \mathcal{A}$ with $c \leq a$ and $c \in l_{\mathcal{N}}(x, y)$.
We have $c \leq a \leq a + b$, so $(x, y) \in \operatorname{rep}(a + b)$.

\item Let us show that $\operatorname{rep}(a ; b) = \operatorname{rep}(a) ; \operatorname{rep}(b)$.

Suppose $(x, y) \in \operatorname{rep}(a ; b)$. We need some $z$ with $(x, z) \in \operatorname{rep}(a)$ and $(z, y) \in \operatorname{rep}(b)$.

@@

By the condition, $\forall$ makes the witness moves and $\exists$ responds with a witness. Her response is a fresh node $z$ such that $l_{\mathcal{N}_{i + 1}}(x, z) =  \up{(l_{\mathcal{N}_i}(x, z) \cup \{ a \})}$ and $l_{\mathcal{N}_{i + 1}}(z, y) = \up{(l_{\mathcal{N}_i}(z, y) \cup \{ b\})}$. The inclusion $\operatorname{rep}(a ; b) \subseteq \operatorname{rep}(a) ; \operatorname{rep}(b)$ holds since all witness moves have been played and for every such a move there exists a witness.

\item Suppose $(x, y) \in \operatorname{rep}(a) ; \operatorname{rep}(b)$. We need $(x, y) \in \operatorname{rep}(a;b)$. That is, there exists $z \in \operatorname{Nodes}(\mathcal{N})$ with $(x, z) \in \operatorname{rep}(a)$ and $(z, y) \in \operatorname{rep}(b)$. So, there are $c, d$ such that $c \leq a$ with $c \in l_{\mathcal{N}}(x, z)$ and $d \leq b$ with $d \in l_{\mathcal{N}}(z, y)$.
We also know that $l_{\mathcal{N}}(x, z) ; l_{\mathcal{N}}(z, y) \subseteq l_{\mathcal{N}}(x, y)$ because all composition moves have been played.
So $c ; d \in l_{\mathcal{N}}(x, y)$. That makes $(x, y) \in \operatorname{rep}(a;b)$ since $c ; d \leq a ; b$.

@@

Let us define the equivalence relation $\approx$ on $T$ as follows for terms $t_1, t_2$:
\begin{center}
$t_1 \approx t_2$ iff $F(\omega) \models t_1 = t_2$.
\end{center}
where $F(\omega)$ is the free countably generated algebra join semilattice-ordered semigroup.
We have the following partial order on $T$ defined as
\begin{center}
  $t_1 \preceq t_2$ iff $t_1 + t_2 \approx t_2$, that is, $F(\omega) \models t_1 + t_2 = t_2$
\end{center}

@@

Let $\mathcal{N}$ be a term network. We define the following three extensions of $\mathcal{N}$ reflecting the composition, witness, and join moves respectively:
\begin{enumerate}
  \item Let $x, y \in \operatorname{Nodes}(\mathcal{N})$ and let $t$ be a term. $\mathcal{N}_c$ is the extension of $\mathcal{N}$,
  where $\operatorname{Nodes}(\mathcal{N}_c) = \operatorname{Nodes}(\mathcal{N})$ and $\up{l_{\mathcal{N}}(x, y) \cup t}$ and $l_{\mathcal{N}_c}(u, v) = l_{\mathcal{N}}(u, v)$ for each $u \neq x$ and $v \neq y$. We denote this network as $\mathcal{N}_c(\mathcal{N}, x, y, t)$.
  \item
  Let $x, y \in \operatorname{Nodes}(\mathcal{N})$, let $z$ be a fresh node, and $t_1$, $t_2$ any terms. Let us define a network $T$ such that $\operatorname{Nodes}(T) = \{ x, y, z\}$. We define labelling as $l_{T}(x, y) = \up{t_1}$ and $l_{T}(y, z) = \up{t_2}$. So we put $\mathcal{N}_w = \mathcal{N} \cup T$.
  We denote this network as $\mathcal{N}_w(\mathcal{N}, x,y,z, t_1, t_2)$.
  \item Let $x, y \in \operatorname{Nodes}(\mathcal{N})$ and let $t_1, t_2$ be terms
  Let us define $T_i = \langle \{ x, y\}, \{ (x, y) \}, l_{T_i} \rangle$, where $l_{T_i}(x, y) = \up{(l_{\mathcal{N}}(x, y) \cup t_i)}$ for $i = 1,2$. So $\mathcal{N}_{j_1} = \mathcal{N} \cup T_1$ and $\mathcal{N}_{j_2} = \mathcal{N} \cup T_2$.
  We denote this as $\mathcal{N}_j(\mathcal{N}, x, y, t_1, t_2)$.
\end{enumerate}

@@

If we extend $\mathcal{A}$, a finite join semilattice-ordered semigroup, with the bottom, then we may consided the obtained algebra as a quantale. By Theorem~\label{quantaleRep}, $\mathcal{A}$ is isomorphic to the relational quantale $\widehat{A}$. From the other hand, if $\mathcal{Q}$ is a relational quantale, then $\Sigma A \subseteq \bigcup A$ for $A \subseteq \mathcal{Q}$. The converse inclusion does not hold in general, see \cite[Remark 3.2 and Example 3.6]{brown1993representation}.

@@

A quantale $\mathcal{Q}$ is called \emph{simple}, if
$\sum \limits_{i \in I} \widehat{A} = \bigcup \limits_{i \in A} \widehat{A}$ for any $A \subseteq \mathcal{Q}$.

@@

Note that an arbitrary finite $\mathcal{A} \in {\bf R}(;,+)$ does not have to be a quantale since a finite join semilattice may not have the bottom. Let us extend $\mathcal{A}$ with the bottom defined as $0 = \Sigma \emptyset$, we denote this algebra as $\mathcal{A}_0$. Note that $a_1 ; a_2 = 0$, where $a_1, a_2 \in \mathcal{A}_0$ and one has either $a_1 = 0$ or $a_2 = 0$.

\begin{proposition}
  If $\mathcal{A} \in {\bf R}(;,+)$ is finite, then $\mathcal{A}_0$ is representable.
\end{proposition}

\begin{proof}
  Let $h$ be a representation of $\mathcal{A}$ over some base set $D$. Let us define a map $h_0 : \mathcal{A}_0 \to 2^{D \times D}$ such that $h_0(a) = h(a)$, if $a \neq 0$, and $h_0(0) = \emptyset$. $h_0$ is a representation since $h$ is a representation.
\end{proof}

\begin{proposition} \label{distr}
  Let $\mathcal{A} \in {\bf R}(;,+)$ and $\mathcal{A}$ is finite, then
  \begin{enumerate}
    \item $\mathcal{A}_0$ is distributive,
    \item Every generator of $\mathcal{A}_0$ as a quantale is join-prime.
  \end{enumerate}
\end{proposition}

@@

  \begin{enumerate}
    \item
    $\mathcal{A}$ is finite, so the $\operatorname{Id}(\mathcal{A}_0)$ consists of principal ideals, that is, lower cones generated by elements of $\mathcal{A}$.

    \begin{center}
    $\operatorname{Id}(\mathcal{A}_0) = \{ \down{a} \: | \: a \in \mathcal{A} \}$
    \end{center}

    Let $h$ be a representation of $\mathcal{A}$, consider $h_0(\mathcal{A}_0)$ and its set of ideals $\operatorname{Id}(h_0(\mathcal{A}_0))$. Intersection of ideals is defined is usual since intersection of ideals is an ideal, join is defined as follows:

    \begin{center}
    $\down{h_0(a)} \vee \down{h_0(b)} = \down{(h_0(a) \cup h_0(b))}$
    \end{center}
    Note that $\down{(h_0(a) \cup h_0(b))} = \down{h_0(a + b)}$ since $h$ is a representation of $\mathcal{A}$.

@@

    Suppose $h(d) \in \down{h_0(c)} \cap \down{h_0(a + b)}$. We need $h_0(d) \in \down{(h_0(c) \cap h_0(a))} \vee \down{(h_0(c) \cap h_0(b))}$.

    Let $(x, y) \in h_0(d)$, then either $(x, y) \in h_0(a)$ or $(x, y) \in h_0(b)$ since $d \leq a + b$.
    Suppose $(x, y) \in h_0(a)$. We need $h_0(d) \in \down{(h_0(c) \cap h_0(a))}$, that is, $h_0(d) \subseteq h_0(c) \cap h_0(a)$, but if $(x, y) \in h_0(d)$, then $(x, y) \in h_0(c)$ since $h_0(d) \subseteq h_0(c)$. Therefore, every $(x, y) \in h_0(d)$ belongs to $h_0(c) \cap h_0(a)$, if $h_0(d)$ is a subset of $h_0(a)$.

    The second alternative with the assumption that $(x, y) \in h_0(a)$ is similar to the previous one.

    Suppose $h_0(d) \in (\down{h_0(c)} \cap \down{h_0(a)}) \vee (\down{h_0(c)} \cap \down{h_0(b)})$. So, we have either $h_0(d) \in \down{h_0(c)} \cap \down{h_0(a)}$ or $h_0(d) \in \down{h_0(c)} \cap \down{h_0(b)}$. So $h_0(d) \in \down{h_0(c)}$ anyway. If $h_0(d) \in \down{h_0(a)}$, then $d \leq a \leq a + b$, so $h_0(d) \in \down{h_0(a + b)}$.
    The second alternative is the same as the first one.

    So, $\operatorname{Id}(h_0(\mathcal{A}))$ is distributive, and, therefore, $\mathcal{A}_0$ is distibutive as a join-semilattice by \cite[Lemma 184]{gratzer2011lattice}.

    \item Let $g \in \mathcal{A}_0$ be a generator, suppose $g \leq a + b$, so there are $a' \leq a$ and $b' \leq b$ with $g = a' + b'$. $g$ is join-irreducible, thus, one has either $g = a'$ or $g = b'$. In fact, this statement generalises the fact that an element is join-prime iff it is join-irreducible in distributive lattices, see \cite[Lemma 5.11]{davey2002introduction}.
  \end{enumerate}
\end{proof}

The following lemma allows one to consider extensions finite elements of ${\bf R}(;,+)$ with bottoms as simple quantales. If $\mathcal{A}$ is finite and representable, then every $a \in \mathcal{A}_0$ has the form $\Sigma \{ g \in \mathcal{G}(\mathcal{A}) \: | \: g \leq a \}$, but every generator is join-prime by Proposition~\ref{distr}. So we have:

\begin{proposition}\label{quantalerepisok}
  Let $\mathcal{A}$ be a finite representable join-semilattice semigroup, then $\mathcal{A}_0$ is completely coprime algebraic as a quantale and, therefore, is simple.
\end{proposition}

@@

\begin{proof}
  Let us show that if $\mathcal{A} \in {\bf R}(+, ;)$ such that $\mathcal{A}$ is finite. Let $h$ be a representation of $\mathcal{A}$, so $h_0$ is a representation of $\mathcal{A}_0$.
  $h(\mathcal{A})$ is a subalgebra of $h_0(\mathcal{A}_0)$, but
  $h_0(\mathcal{A}_0)$ is isomorphic to $\widehat{A_0}$, the relational quantale representation of $\widehat{A_0}$ for the following reason.

  Consider the map
  $f : h_0(a) \mapsto \hat{a}$. $f$ preserves composition since $h$ and $\hat{.}$ are semigroup homomorphisms. The equality $f(h_0(a) \cup h_0(b)) = \hat{a} \cup \hat{b}$ holds since $\mathcal{A}_0$ is simple as a quantale by Proposition~\ref{quantalerepisok}. $f$ is also obviously injective and surjective.

  $\widehat{\mathcal{A}_0}$ is a relational quantale on the underlying set of $\mathcal{A}_0$, but $\mathcal{A}$ is finite, so $\mathcal{A}$ has a representation over a finite base since $h(\mathcal{A})$ is isomorphic to $h_0(\mathcal{A})$, a subalgebra of $\widehat{A_0}$.
\end{proof}

@@

The author is sincerely grateful to Robin Hirsch, Ian Hodkinson, and his supervisor Ilya Shapirovsky for valuable comments. The previous version of this paper contained an error. Many thanks to Ja\v{s} \v{S}emrl for noticing that.
